Using physics simulations to predict children's and adults' preference for physical interactions

### 2021-02-04 paper/paper2021.Rmd
GD edits

Given the high 
<!-- split-half -->
reliability of choices on this task, particularly for adults ($r = .95$, children's $r = .61$), we tested how well heuristic models of curiosity based on simulations of the physical interactions could predict participants' preferences of target objects on each trial.
We tested a wide range of features that we thought might be proportional to people's interest in particular object interactions (e.g., how far a dropped object might end up from the drop location), but the feature that best predicted both children's and adults' responses was the probability that a dropped object would come to rest atop a target object (*Pr(support)*). <!-- Adults were best fit by a softmax $\beta = 1.18$, while children were best modeled by $\beta = 0.5$, meaning that as far as this model is concerned children are "noisy" adults. -->
However, on a trial level this model's choices were negatively correlated with people's. 
One of the next best-fitting models was the sensitivity of *Pr(support)* to perturbations in the initial starting conditions, which broadly was able to capture people's stronger preference in *contain* condition trials relative to *support* condition trials, but still did not capture item-level variation in people's choices within condition.
On balance, although much of the variation in this dataset is unexplained, it is encouraging that people's interest in physical interactions can be in part accounted for by simple heuristic models based on simulations of the same physical task we asked participants to merely imagine.

At the same time, the fact that these results are not fully explainable by simple heuristic models suggest that this paradigm may fruitfully serve as a test bed for evaluating more complex computational theories of curiosity. For example, in one proposal learners prefer to explore stimuli that are "moderately discrepant" in relation to their current knowledge state, thereby providing an opportunity to learn [@Kinney1976], and curiosity changes as the gap between the learner's knowledge develops and the state of the world closes [@Loewenstein1994].
These theories are starting to be implemented in deep neural networks, which are now capable of learning forward and inverse physical dynamics (i.e., "intuitive physics") from images when given the ability to "poke" the objects in the scene [@Agrawal2016], and have more recently been used to test which "curious" policies for generating actions result in robust and effective learning of intuitive physics in deep reinforcement learning agents [@Haber2018learning].
<!-- @Haber2018learning created a deep reinforcement learning model to drive intrinsically-motivated artificial agents, using two interacting neural networks: a *world-model* and a *self-model*.
These networks are trained simultaneously to solve different problems: the world-model attempts to predict the consequences of the agent's actions (either the forward or inverse dynamics), while the self-model attempts to learn to predict the errors of the world-model. 
The agent can use the predictions of the self-model to implement a variety of curiosity policies for choosing actions, for example to pick actions that it believes will challenge its world-model. 
-->
In future work, we aim to examine the development of curiosity in such curious artificial agents, and compare these agents' behavior to people's in the same tasks.
Curiosity has inspired curiosity in researchers for decades, but only recently have the tools come along that make these theories testable. We propose that the development of parallel tasks for children, humans, and embodied agents is a promising way forward towards modeling curiosity in the real-world.

>>>>

Given the high reliability of choices on this task, particularly for adults ($r = .95$, children $r = .61$), we tested how well heuristic models of curiosity based on simulations of the physical interactions could predict participants' preferences of target objects on each trial.
We tested a wide range of features that we thought might be proportional to people's interest in particular object interactions (e.g., how far a dropped object might end up from the drop location), but not one of these features explained much of the trial-level variance in people's choices. 
Instead, we found a few features that were partially correlated with people's choices: for example, the inverse standard deviation of the target's final position was the best predictor of both children's ($r=.41$) and adults' choices ($r=.41$) on support trials.
For containment trials, the normed velocity of the drop object after first collision was most correlated with adults' choices ($r=.54$) and moderately with children's ($r=.41$), but children's choices were also correlated with the maximum and mean distance traveled by the target ($r=.43$ and $r=.42$).
On balance, although much systematic variation in this dataset is unaccounted for, it is somewhat encouraging that people's interest in physical interactions can be in part accounted for by simple heuristic models based on simulations of the same physical task we asked participants to merely imagine.

At the same time, the fact that these results are not fully explainable by simple heuristic models suggest that this paradigm of "dropping it like it's hot" may fruitfully serve as a test bed for evaluating more complex computational theories of curiosity. 
For example, in one proposal learners prefer to explore stimuli that are "moderately discrepant" in relation to their current knowledge state, thereby providing an opportunity to learn [@Kinney1976], and curiosity changes as the gap between the learner's knowledge develops and the state of the world closes [@Loewenstein1994].
These theories are beginning to be implemented in deep neural networks, which are now capable of learning forward and inverse physical dynamics (i.e., "intuitive physics") from images when given the ability to "poke" the objects in the scene [@Agrawal2016], and have more recently been used to test which "curious" policies for generating actions result in robust and effective learning of intuitive physics in deep reinforcement learning agents [@Haber2018learning].
<!-- @Haber2018learning created a deep reinforcement learning model to drive intrinsically-motivated artificial agents, using two interacting neural networks: a *world-model* and a *self-model*.
These networks are trained simultaneously to solve different problems: the world-model attempts to predict the consequences of the agent's actions (either the forward or inverse dynamics), while the self-model attempts to learn to predict the errors of the world-model. 
The agent can use the predictions of the self-model to implement a variety of curiosity policies for choosing actions, for example to pick actions that it believes will challenge its world-model. 
-->
In future work, we aim to examine the development of curiosity in such curious artificial agents, and compare these agents' behavior to people's in the same tasks.
The multifaceted nature of curiosity has fascinated researchers for decades, but only recently have we developed the means to factorize and test these theories. 
We propose that the development of parallel tasks for children, humans, and embodied agents is a promising way forward towards modeling curiosity in the real-world.
